---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"   
)
```

# lifeRstyle

<!-- badges: start -->
<!-- badges: end -->

## Description

lifeRstyle is an R package designed to download, clean, and analyse
Irish lifestyle and health data from the Central Statistics Office (CSO).
The package focuses on indicators such as:

Smoking habits

Alcohol consumption

General health and wellbeing

It provides a tidy, user-friendly workflow for loading data, visualising trends,
and fitting statistical models.

Three core functions are provided:

1. Load_LifeRStyle() – for downloading and preparing CSO lifestyle data

2. plot() – for producing visualisations using ggplot2

3. fit() – for fitting statistical models (linear regression, ANOVA, mixed-effects models)

Typically, these functions are applied sequentially.
A demonstration of the typical workflow is shown below.


## Installation

Install the development version of lifeRstyle from [GitHub](https://github.com/) with:

``` {r, eval=FALSE}
# install.packages("pak")
pak::pak("DavidF2021/lifeRstyle")

```
## Example

We are showing how to load lifestyle data and perform a simple analysis:

```{r}
library(lifeRstyle)

data <- lifeRstyle::combined_cleaned
```
Once loaded, the data can be visualized easily:
```{r plot, warning=FALSE}
suppressWarnings(
  plot_combined_data(data)
)

```
Statistical models
```{r lm}
model <- fit_lifeRstyle(data, fit_type = "lm")
summary(model) 

```

Anova model
```{r anova}
model_anova <- fit_lifeRstyle(data, fit_type = "anova")
summary(model_anova)
```
Mixed models
```{r mixed}
model_mixed <- fit_lifeRstyle(data, fit_type = "mixed")
summary(model_mixed)

```
